---
title: "Project Draft"
author: "Haley Vien"
date: "12/3/2019"
output: html_document
df_print: paged
---

```{r setup, include=FALSE}
library(fivethirtyeight)
library(ggplot2)
data("hiphop_cand_lyrics")
library(broom)
library(moderndive)
library(tidyverse)
library(openintro)
library(ggridges)
```

# The Data

The data we will be using is *hiphop_cand_lyrics* from the *fivethirtyeight* package. This data represents when hiphop songs referenced POTUS candidates in 2016. There are 377 rows of data and 8 variables. The 8 variables are the *candidate* referenced, the *song* name, the *artist*, the *sentiment* of the lyric, the *theme* of the lyric, the *album_release_date*, the *line* referencing the candidate, and the *url* to the song. This data was collected from [Genius](https://genius.com/).

Note: For themes, one of the values is NA which represents misc. themes.

Note 2: The references were classified as negative, positive, and neutral by a group of FiveThirtyEight staffers who were interested in hip-hop and quantifying things. If there was a reference that obviously fit into one category, the lead researcher, Allison McCann, would classify it. Additionally, if there was a a trickier reference that was hard to decipher, three people, senior editor Chadwick Matlin, intern Setrige Crawford and McCann, would make the final decision after deliberation. 

# Research Question

When looking at Hillary Clinton and Donald Trump’s mentions in hip hop songs, does the relation between candidate or theme and sentiment differ between the two and does that change over time?

Our response variable is sentiment while our explanatory variables are candidate, theme, and the album's release date. 

# Graphing the Data

```{r, echo=FALSE}
hiphop_hcdt <- hiphop_cand_lyrics %>% 
  filter(candidate %in% c("Hillary Clinton", "Donald Trump")) %>% 
  mutate(sentiment_negative = ifelse(sentiment == "negative", 1, 0))
```


```{r, echo=FALSE}
hiphop_hc <- hiphop_cand_lyrics %>% 
  filter(candidate %in% c("Hillary Clinton"))
```

```{r, echo=FALSE}
hiphop_dt <- hiphop_cand_lyrics %>% 
  filter(candidate %in% c("Donald Trump"))
```

```{r, echo=FALSE}
ggplot(data = hiphop_hcdt) + 
  geom_bar(aes(x = sentiment, fill = theme)) +
  facet_wrap(vars(candidate))
```

The graph above shows the sentiment of the songs that mention Donald Trump or Hillary Clinton and is filled with the theme of the mentioned lyric. 

Interestingly enough, when analyzing the positive bar for Donald Trump, most of the lyrics that mention him in a positive *sentiment* are related to Trump's hotel and money. The amount of times artists refer to these 2 themes almost quadruples the amount of times Trump is referred to negatively. When looking at the negative bar, most of the negative references are related to the political theme. 

In comparison to Donald Trump, Hillary Clinton has far less references in hiphop songs in total. Interestingly enough, for the positive and neutral bars, the theme that is most prevalent is personal, however in the negative column, the most prevalent theme is political. Even though there are slight differences, I thought it was interesting that Clinton in general had more references to the sexual theme than Trump. 

```{r, echo=FALSE}
ggplot(data = hiphop_hcdt)+
   geom_density_ridges(aes(x = album_release_date, fill = candidate, y = sentiment), alpha = .5)

```


The above graph showcases how the *sentiment* towards Donald Trump and Hillary Clinton have changed over time in reference to the *album_release_date* of songs that reference them.

Between 2010 and 2020, the negative *sentiment* towards Donald Trump drastically increases and even peaks at around 2016, right around when he was running for presidency. For Clinton, negative references to her in hiphop songs suddenly increased in 2005 and peaked right before 2010, then it linearly decreased, almost having no negative references in 2016. 

However, when looking at the positive *sentiment*, positive references to Trump peak a few years before 2015 and then decline. However, Clinton begins to have more positive references begining around 2005 and this peaks around 2010 and consistently stays at that peak until around 2016.  

# Fitting models 

```{r, echo=FALSE}
glm.cand <- glm(sentiment_negative ~ candidate, 
                data = hiphop_hcdt, 
                family = binomial)

tidy(glm.cand)
```

*Intercept*: Log-odds of positive being the sentiment for Donald Trump is 1.93.

*candidateHillaryClinton*: difference in log-odds of positive being the sentiment between Donald Trump and Hillary Clinton 


```{r, echo=FALSE}
glm.cand.date <- glm(sentiment_negative ~ candidate + album_release_date, 
                     data = hiphop_hcdt, 
                     family = binomial)

tidy(glm.cand.date)
```

*Intercept*: Log-odds of positive being the sentiment for Donald Trump is 120.63.

*candidateHillaryClinton*: Difference in log-odds of positive being the sentiment between Hillary Clinton and Donald Trump. 

*album_release_date*: Difference in log-odds of positive being the sentiment between Hillary Clinton and Donald Trump accounting for *album_release_date*.

# Confidence Intervals

```{r, echo=FALSE,message=FALSE, warning=FALSE}
tidy(glm.cand) %>% 
  mutate(lowerCI = estimate - 2*std.error,
         upperCI = estimate + 2*std.error)

```

This tells us that the plausible range of values for the true difference in sentiment. Since it is between -1.73 and -0.57, the difference in sentiment between Donald Trump and Hillary Clinton appears to be fairly large. 

```{r, echo=FALSE}
tidy(glm.cand.date) %>% 
  mutate(lowerCI = estimate - 2 * std.error,
         upperCI = estimate + 2*std.error)
```

This tells us that the plausible range of values for the true difference in sentiment. Since it is between -0.1093439 and -0.0088, the difference in sentiment between Donald Trump and Hillary Clinton accounting for *album_release_date* appears to be even larger. 

# Next steps
(@) Summarize our results
(@) Ask for help and attend office hours to fix mistakes
(@) Schedule a time to meet to discuss presentation details
