---
title: "Project Draft"
author: "Haley Vien"
date: "12/3/2019"
output: html_document
---

```{r setup, include=FALSE}
library(fivethirtyeight)
library(ggplot2)
data("hiphop_cand_lyrics")
library(broom)
library(moderndive)
library(tidyverse)
library(openintro)
library(ggridges)
```

# The Data

The data we will be using is *hiphop_cand_lyrics* from the *fivethirtyeight* package. This data represents when hiphop songs referenced POTUS candidates in 2016. There are 377 rows of data and 8 variables. The 8 variables are the *candidate* referenced, the *song* name, the *artist*, the *sentiment* of the lyric, the *theme* of the lyric, the *album_release_date*, the *line* referencing the candidate, and the *url* to the song. This data was collected from [Genius](https://genius.com/).

Note: For themes, one of the values is NA which represents misc. themes.

# Research Question

When looking at Hillary and Trump’s mentions in hip hop songs, does the relation between candidate or theme and sentiment differ between the two and does that change over time?

Our response variable is sentiment while our explanatory variables are candidate, theme, and the album's release date. 

# Graphing the Data

```{r}
hiphop_hcdt <- hiphop_cand_lyrics %>% 
  filter(candidate %in% c("Hillary Clinton", "Donald Trump"))
```

```{r}
hiphop_hc <- hiphop_cand_lyrics %>% 
  filter(candidate %in% c("Hillary Clinton"))
```

```{r}
hiphop_dt <- hiphop_cand_lyrics %>% 
  filter(candidate %in% c("Donald Trump"))
```



```{r}
ggplot(data = hiphop_hcdt) + 
  geom_bar(aes(x = sentiment, fill = theme)) +
  facet_wrap(vars(candidate))
```

The graph above shows the sentiment of the songs that mention Donald Trump or Hillary Clinton and is filled with the theme of the mentioned lyric. 

Interestingly enough, when analyzing the positive bar for Donald Trump, most of the lyrics that mention him in a positive *sentiment* are related to Trump's hotel and money. The amount of times artists refer to these 2 themes almost quadruples the amount of times Trump is referred to negatively. When looking at the negative bar, most of the negative references are related to the political theme. 

In comparison to Donald Trump, Hillary Clinton has far less references in hiphop songs in total. Interestingly enough, for the positive and neutral bars, the theme that is most prevalent is personal, however in the negative column, the most prevalent theme is political. Even though there are slight differences, I thought it was interesting that Clinton in general had more references to the sexual theme than Trump. 

```{r}
ggplot(data = hiphop_hcdt)+
   geom_density_ridges(aes(x = album_release_date, fill = candidate, y = sentiment), alpha = .5)

```


The above graph showcases how the *sentiment* towards Donald Trump and Hillary Clinton have changed over time in reference to the *album_release_date* of songs that reference them.

Between 2010 and 2020, the negative *sentiment* towards Donald Trump drastically increases and even peaks at around 2016, right around when he was running for presidency. For Clinton, negative references to her in hiphop songs suddenly increased in 2005 and peaked right before 2010, then it linearly decreased, almost having no negative references in 2016. 

However, when looking at the positive *sentiment*, positive references to Trump peak a few years before 2015 and then decline. However, Clinton begins to have more positive references begining around 2005 and this peaks around 2010 and consistently stays at that peak until around 2016.  

<<<<<<< HEAD

# Fitting models 
=======
>>>>>>> dfbbe21b640520b1105399401224dbeec000d55c

```{r}
glm.cand <- glm(factor(sentiment) ~ candidate,
            data = hiphop_hcdt,
           family = binomial(link = "logit"))

tidy(glm.cand)
```


```{r}
glm.cand.date <- glm(factor(sentiment) ~ candidate + album_release_date,
            data = hiphop_hcdt,
           family = binomial(link = "logit"))

tidy(glm.cand.date)
```

# Confidence Intervals




